---
title: "Yield-Curves"
author: "Koudadim Olivier TRAORE"
date: "2025-02-21"
output: ioslides_presentation
---

## The Nelson-Siegel method functional

The Nelson-Siegel method consist of modeling the zero-coupon yield curve using the following parsimonious functional:

$$
R(0, \tau) = \beta_0 + \beta_1 \left( \dfrac{1- \exp(-\tau / \lambda)}{\tau / \lambda}  \right) + \beta_2 \left( \dfrac{1- \exp(-\tau / \alpha)}{\tau / \lambda} -\exp(-\tau / \lambda) \right) 
$$ where:

-   $R(0, \tau)$ denotes the ZC rate at maturity $T$
-   $\lambda$ is the scale parameter
-   $\beta_0$ is the level factor, i.e long-term rate
-   $\beta_1$ is the rotation factor, i.e the distance between the short-term rate and the long-term one
-   $\beta_2$ is the slope factor, measuring the yield curve curvature.

## Nelson-Siegel method's parameters estimation

```{r Labrary}
#install.packages("YieldCurve")
library(YieldCurve)
```

```{r}
data("FedYieldCurve") # Data that will be used
View(FedYieldCurve)
```

```{r}
maturity.Fed <- c(3/12, 0.5, 1, 2, 3, 5, 7, 10)
NSParameters <- Nelson.Siegel( rate=first(FedYieldCurve,'10 month'),	maturity = maturity.Fed)
```

```{r}
NSParameters

y <- NSrates(NSParameters[5,], maturity.Fed)
```

## Plot of Nielson-Siegel Yield Curve

```{r pressure}
plot(maturity.Fed, FedYieldCurve[5,], main="Fitting Nelson-Siegel yield curve", xlab=c("Pillars in months"), type="o")
lines(maturity.Fed, y, col=2)
legend("topleft",legend=c("observed yield curve","fitted yield curve"),
col=c(1,2),lty=1)
grid()
```

## Nelson-Siegel augmented also called Nelson-Siegel-Svenson

The Nelson-Siegel model cannot reconstruct all the yield curves encountered on the market. In particular, the one-hump, one-trough shapes cannot be reproduced. The augmented Nelson-Siegel model remedies this. The functional of the Nelson-Siegel augmented model is given by par :

$$
\begin{aligned}
R(0, \tau) = & \beta_0 + \beta_1 \left( \dfrac{1- \exp(-\tau / \lambda_1)}{\tau / \lambda_1}  \right) + \beta_2 \left( \dfrac{1- \exp(-\tau / \lambda_1)}{\tau / \lambda_1} -\exp(-\tau / \lambda_1) \right) \\
& + \beta_3 \left( \dfrac{1- \exp(-\tau / \lambda_2)}{\tau / \lambda_2} -\exp(-\tau / \lambda_2) \right)
\end{aligned}
$$ where:

-   $R(0, \tau)$ denotes the ZC rate at maturity $T$
-   $\lambda_1$ is the scale parameter
-   $\lambda_2$ is the shape parameter
-   $\beta_0$ is the level factor, i.e long-term rate
-   $\beta_1$ is the rotation factor, i.e the distance between the short-term rate and the long-term one
-   $\beta_2$ is the slope factor (medium-term rate), measuring the yield curve curvature.
-   $\beta_3$ is another slope factor, measuring the yield curve additional curvature.

## Estimation of Nelson-Siegel-Svensson parameters

```{r}
data(ECBYieldCurve)
maturity.ECB <- c(0.25,0.5,seq(1,30,by=1))
A <- Svensson(ECBYieldCurve[1:10,], maturity.ECB )
Svensson.rate <- Srates( A, maturity.ECB, "Spot" )

A
```

```{r}
plot(maturity.ECB, Svensson.rate[5,],main="Fitting Svensson yield curve",
 xlab=c("Pillars in years"), type="l", col=3)
lines( maturity.ECB, ECBYieldCurve[5,],col=2)
legend("topleft",legend=c("fitted yield curve","observed yield curve"),
col=c(3,2),lty=1)
grid()
```
